---
title: 'Week 5 Code-Along: Penguin Size'
author: "Kerry Nixon"
date: "11/15/2020"
output: html_document
---

```{r setup, include=FALSE, echo= TRUE, message= FALSE, warning= FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests
Using t-tests to compare means between groups to see if we have enough evidence to think we drew samples from populations from different means (rejecting the null hypothesis that they were drawn from the same population aka same means and observed differences are sampling variation by chance).

Mass comparision: male & females chinstrap penguins

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

# Exploratory Histograms
ggplot(data = chinstraps, aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

```

Both graphs look somewhat normally distributed, fairly symmetric though not perfect. The roughly normal distribution makes using the mean central tendency a reasonable value for comparison. The male mean appears as though it will be a bit higher than the female mean, with male measure of central tendency closer to 4000g for male and 3500g for female. 

```{r}
# Exploratory QQ-plots to check for normality (more linear = more normal distribution)

ggplot(data = chinstraps, aes(sample = body_mass_g)) +
  geom_qq() +
  facet_wrap(~sex)

```
Both of these look very close to normal distributions as they are quite linear in appearance. Between these plots and the histograms we can feel confident in the normal distribution of our data/samples. 

```{r}
## Get descriptive statistics
chinstraps %>% 
  group_by(sex) %>% 
  summarise(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n())

```

##### Get vectors of body_mass_g for femail and male chinstrap penguins separately for use in t-test.

```{r}

chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```

### T-tests, a couple of different ways

1. Is there a significant difference in body mass between male and female chinstrap penguins?

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest
```
## Welch Two Sample t-test results

1. Not using the assummption of equal variances
2. Lowest line shows the means are given for both samples in the order they were input - should match your summary stats.
3. Alternative hypothesis is given, and you can deduce the null from that - in this case teh null is that the means are the same (the difference is equal to 0) and the alternative is that the difference in means in not equal to 0 (aka the samples are drawn from populations with different means)
4. The "t" value is a measure of how different from 0 the means are from 0 on the t distribution when taking into account sample size and dispersion
5. The "df" is the "degree of freedom" measuring sample size
6. The "p" is "p-value" which measures the probability that we took samples from teh same population (populations with the same mean) with means at least as different as the means we've found in our samples, by random chance.
7. CONTEXT: To add context to the results talk about the:
  . actual difference
  . percent difference
  . effect size (Cohen's D)

```{r}
# Effect Size
cohen.d(chinstraps_f, chinstraps_m)
```

### Using broom::tidy for reproducable writing of results.

There is a significant difference in body mass, and a large effect size (Cohen's *d* = -1.26). The way I'm *writing this* is not reproducible. How can we use in-line referencing of outcomes so that the values would automatically update if the values did? 

Use `broom::tidy()` to get messy model outputs into tidy data frames that are easy to reference! 

```{r}
chinstrap_out <- broom::tidy(chinstrap_ttest)
chinstrap_out
```

You can **reference values from model outputs in your R markdown text**:

t(`r round(chinstrap_out$parameter,2)`) = `r round(chinstrap_out$statistic,2)`, *p* = `r chinstrap_out$p.value`.

female chinstrap mean = `r round(chinstrap_out$estimate1)`g

male chinstrap mean = `r round(chinstrap_out$estimate2)`g

##### One-sided t-test

Are male chinstrap penguins significantly larger (body mass) than female chinstrap penguins?

```{r}
# Updating the alternative hypothesis to see whether the mean of first variable - chinstraps_m (males) are greater than second variable, chinstraps_f (female)

t.test(chinstraps_m, chinstraps_f, alternative = "greater") 

# Because this is only looking at one side of the difference (greater than rather than different from which includes both greater than and less than), the upper tail only, it will have half of the p-value as the two-sided t-test
```
# Part 2. Intro to ANOVA and Beeswarm Graphs

### One-way ANOVA

###### 1. Table of counts by species and sex

```{r}
penguin_counts <- penguins %>% 
  count(species, sex)

penguin_counts
```

###### 2. Create a subset of only female penguins and data exploration

```{r}
# Exploratory graphs and stats for female penguins and flipper length
penguin_f <- penguins %>% 
  filter(sex == "female")

# Histograms
ggplot(data = penguin_f, aes(x = flipper_length_mm)) +
  geom_histogram(aes(fill=species),
                 show.legend = FALSE) +
  facet_wrap(~species,
             scales = "free")
# QQ-plot
ggplot(data = penguin_f, aes(sample = flipper_length_mm)) +
  geom_qq(aes(color=species),
                 show.legend = FALSE) +
  facet_wrap(~species,
             scales = "free")
```

###### 3. Summary Statistics

```{r}
penguin_f_summary <- penguin_f %>% 
  group_by(species) %>% 
  summarize(
    mean_flipper_length = mean(flipper_length_mm),
    sd_flipper_length = sd(flipper_length_mm),
    sample_size = n(),
    se_flipper_length = sd(flipper_length_mm)/sqrt(n()),
    variance_flipper_length = var(flipper_length_mm)
  )

# Make sure to check if the largest variance is within 4X of the smallest variance to see if data qualify for ANOVA test - for this data its close to 2X which is well within the acceptable range.

# Using the Levene Variance test - use with caution with respect to sample size as large sample size can cause a significant variance, and refer to the actual variances to check for the 4X range. (same 0.05 significance level)

leveneTest(flipper_length_mm ~ species, penguin_f)

penguin_f_summary # returning the summary table
```
###### 4. Beeswarm Plot with Mean +/- 1 sd

```{r}
ggplot() + 
  geom_beeswarm(data = penguin_f,
                aes(x = species, y = flipper_length_mm),
                color = "gray",
                size = 2,
                alpha = 0.5) +
  geom_point(data = penguin_f_summary, 
             aes(x = species, y = mean_flipper_length),
             size = 4) +
  geom_errorbar(data = penguin_f_summary,
                aes(x = species,
                    ymax = mean_flipper_length + sd_flipper_length,
                    ymin = mean_flipper_length - sd_flipper_length),
                width = 0.1) +
  theme_minimal()
```

###### 5. One-way ANOVA

```{r}

# Usually you want type 3 anova due to unequal sample sizes etc

# Type 1 ANOVA
penguin_aov <- aov(flipper_length_mm ~ species, data = penguin_f)
summary(penguin_aov)
```
The "F value" is the ration of the between group variance divided by the within group variance. The larger the F stat that means there is a larger between group variance. Larger between group variance, and therefore larger F Value will be reflected in a smaller p-value. A smaller p-value indicates a lower probability that the smaples were drawn from a population with the same mean and have means at least as different as these means, due to randome chance. With this example we would reject the null hypothesis and would now perform some post-hoc tests.

```{r}
# Post-hoc test for pair-wise comparisons: TukeyHSD

TukeyHSD(penguin_aov)
```
All of the values are below 0.05 and so we would say there is a significant size difference between all of the groups. 

###### Type III ANOVA using car package

```{r}
# If you have very unbalanced design, with large differences in samples sizes etc, you will use an orthogonal or type 3 ANOVA with the car package
penguin_lm <- lm(flipper_length_mm ~ species, data = penguin_f)
penguin_aov_t3 <- Anova(penguin_lm, type = "III")
penguin_aov_t3
```

I have now run t-tests, done tons of exploratory graphing and run two types of ANOVAs! 

